---
title: "How to use shapper for regression"
author: "Alicja Gosiewska"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{How to use shapper for regression}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# Introduction

The `shapper` is an R package which ports the `shap` python library in R. For details and examples see [shapper repository on github](https://github.com/ModelOriented/shapper) and [shapper website](https://modeloriented.github.io/shapper/).

SHAP (SHapley Additive exPlanations) is a method to explain predictions of any machine learning model. 
For more details about this method see [shap repository on github](https://github.com/slundberg/shap).

# Install shaper and shap

## R package shapper

```{r, eval = FALSE}
devtools::install_github("ModelOriented/shapper")
```

## Python library shap

To run shapper python library shap is required. It can be installed both by python or R. To install it throught R, you an use function `install_shap` from the `shapper` package.

```{r, eval = FALSE}
shapper::install_shap()
```


# Load data sets

The example usage is presented on the `HR` dataset from the R package `DALEX2`. For more details see [DALEX2 github repository](https://github.com/ModelOriented/DALEX2).

```{r}
# devtools::install_github("ModelOriented/DALEX2")
library("DALEX2")
Y_train <- dragons$life_length
x_train <- dragons[ , -8]
```


# Let's build model

```{r}
library(randomForest)
set.seed(123)
model_rf <- randomForest(x = x_train, y = Y_train)
```


# Here shapper starts

First step is to create an explainer for each model. The explainer is an object that wraps up a model and meta-data.

```{r}
library(shapper)

ive_rf <- individual_variable_effect(model_rf, data = x_train,
                                     new_observation = x_train[1,])
```

```{r}
ive_rf
```

# Plotting results

```{r}
plot(ive_rf)
```







